---
title: "R Notebook"
output: html_notebook
---

Problem statement:

Loading all required library
```{r}
library(moments) #skewness
library(pastecs) #descriptive statistics stats.desc
library(diptest) # Hartigans' dip test for unimodality 
library(fmsb) #VIF
library(DAAG) #K-fold validation
library(ggplot2)
```

Data import
Initial setwd command will set current working directory. The file in the format of csv(comma separated version) loaded to R in data.frame format.
str command will give observation,data type, name, few sample of variables.

Attribute Information:


```{r}
# data mining
setwd("/home/sudhir/git/ML-R/OLS project/Cement")
data=read.csv('train.csv')
str(data)
head(data)
```

Finding missing value "NA" using "is.na()".
```{r}
# Checking missing value
summary(is.na(data)) # Column wise summary
table(is.na(data)) # Total summary
```

Measure of central tendency of data.
```{r}
summary(data)# M.o.CT
stat.desc(data)# M.o.CT +M.o.V
```

A boxplot provides a graphical view of the median, quartiles, maximum, and minimum of a data set. Outlier is an observation point that is distant from other observations.

A histogram is a graphical representation of the distribution of numerical data. It is an estimate of the probability distribution of a continuous variable.To contruct histogram,the first step is to "bin" the range of values—that is, divide the entire range of values into a series of intervals—and then count how many values fall into each interval.

plot(density(data$...) will give distribution of data

Univariate analysis of "Cement" attribute
```{r}
new=data
boxplot(new$Cement, ylab='Cement(Kg)', main='Cement Boxpolt',col = 'red' )
hist(new$Cement, xlab='Cement(Kg)', main= 'Cement histogram', col='green')
plot(density(new$Cement), xlab='Cement(Kg)', main= 'Cement Density plot', col='blue')
skewness(new$Cement)
dip.test(new$Cement)
```

Univariate analysis of "Slag(Kg)" attribute
```{r}
boxplot( new$Slag, ylab='Slag(Kg)', main='Slag Boxpolt',col = 'red' )
hist(new$Slag, xlab='Slag(Kg)', main= 'Slag histogram', col='blue')
plot(density(new$Slag), xlab='Slag(Kg)', main= 'Slag Density plot', col='darkgreen')
skewness(new$Slag)
dip.test(new$Slag)
```

Univariate analysis of "Ash(Kg)" attribute
```{r}
boxplot( new$Ash, ylab='Ash(Kg)', main='Ash Boxpolt',col = 'green' )
hist(new$Ash, xlab='Ash(Kg)', main= 'Ash histogram', col='red')
plot(density(new$Ash), xlab='Ash(Kg)', main= 'Ash Density plot', col='darkblue')
skewness(new$Ash)
dip.test(new$Ash)
```

Univariate analysis of "Water(Kg)" attribute
```{r}
boxplot( new$Water, ylab='Water(Kg)', main='Water Boxpolt',col = 'yellow' )
hist(new$Water, xlab='Water(Kg)', main= 'Water histogram', col='blue')
plot(density(new$Water), xlab='Water(Kg)', main= 'Water Density plot', col='purple')
skewness(new$Water)
dip.test(new$Water)
```

Univariate analysis of "Superplasticizer(Kg)" attribute
```{r}
boxplot( new$Superplasticizer, ylab='Superplasticizer(Kg)', main='Superplasticizer Boxpolt',col = 'green' )
hist(new$Superplasticizer, xlab='Superplasticizer(Kg)', main= 'Superplasticizer histogram', col='brown')
plot(density(new$Superplasticizer), xlab='Superplasticizer(Kg)', main= 'Superplasticizer Density plot', col='blue')
skewness(new$Superplasticizer)
dip.test(new$Superplasticizer)
```

Univariate analysis of "Coarse Aggregate(Kg)" attribute
```{r}
boxplot( new$CoarseAggregate, ylab='CoarseAggregate(Kg)', main='CoarseAggregate Boxpolt',col = 'red' )
hist(new$CoarseAggregate, xlab='CoarseAggregate(Kg)', main= 'CoarseAggregate histogram', col='green')
plot(density(new$CoarseAggregate), xlab='CoarseAggregate(Kg)', main= 'CoarseAggregate Density plot', col='yellow')
skewness(new$CoarseAggregate)
dip.test(new$CoarseAggregate)
```

Univariate analysis of "Fine Aggregate(Kg)" attribute
```{r}
boxplot( new$FineAggregate, ylab='FineAggregate(Kg)', main='FineAggregate Boxpolt',col = 'red' )
hist(new$FineAggregate, xlab='FineAggregate(Kg)', main= 'FineAggregate histogram', col='yellow')
plot(density(new$FineAggregate), xlab='FineAggregate(Kg)', main= 'FineAggregate Density plot', col='blue')
skewness(new$FineAggregate)
dip.test(new$FineAggregate)
```

Univariate analysis of "CompressiveStrength" attribute
```{r}
boxplot( new$CompressiveStrength, ylab='CompressiveStrength(Kg)', main='CompressiveStrength Boxpolt',col = 'red' )
hist(new$CompressiveStrength, xlab='CompressiveStrength(Kg)', main= 'CompressiveStrength histogram', col='green')
plot(density(new$CompressiveStrength), xlab='CompressiveStrength(Kg)', main= 'CompressiveStrength Density plot', col='red')
skewness(new$CompressiveStrength)
dip.test(new$CompressiveStrength)
```

Transformed output applying BOX COX tranfermation
```{r}
#library('caret')
c<- caret:: BoxCoxTrans(new$CompressiveStrength)
print(c)
new$CompressiveStrength1=new$CompressiveStrength^0.5
```

Bivaraite analysis
The covariance matrix generalizes the notion of variance to multiple dimensions.
A scatter plot provides a graphical view of the relationship between two sets of attributes.
Created new variable for Compressive Strength for Days (is factor) by mean. 
```{r}
cor(new[,-8])
plot(new$Cement,new$CompressiveStrength1,col=c('red','green'),pch=16, xlab = 'Cement in (kG)',ylab = 'CompressiveStrength1 (Mpa)', main = 'Cement vs CompressiveStrength1')
plot(new$Slag,new$CompressiveStrength1,col=c('blue','darkgreen'),pch=16, xlab = 'Slag in (kG)',ylab = 'CompressiveStrength (Mpa)', main = 'Slag vs CompressiveStrength')
plot(new$Ash,new$CompressiveStrength1,col=c('red','green'),pch=16, xlab = 'Ash in (kG)',ylab = 'CompressiveStrength (Mpa)', main = 'Ash vs CompressiveStrength')
plot(new$Water,new$CompressiveStrength1,col=c('blue','brown'),pch=16, xlab = 'Water in (kG)',ylab = 'CompressiveStrength (Mpa)', main = 'Water vs CompressiveStrength')
plot(new$Superplasticizer,new$CompressiveStrength1,col=c('red','green'),pch=16, xlab = 'Superplasticizer in (kG)',ylab = 'CompressiveStrength (Mpa)', main = 'Superplasticizer vs CompressiveStrength')
plot(new$CoarseAggregate,new$CompressiveStrength1,col=c('blue','darkgreen'),pch=16, xlab = 'CoarseAggregate in (kG)',ylab = 'CompressiveStrength (Mpa)', main = 'CoarseAggregate vs CompressiveStrength')
plot(new$FineAggregate,new$CompressiveStrength1, col=c('red','purple'),pch=16, xlab='FineAggregate', ylab = 'CompressiveStrength',main='FineAggregate vs CompressiveStrength')

ag=aggregate(new$CompressiveStrength1, by=list(new$Days),mean)
ag
pairs(new[,-8],col=c('blue','yellow'))
```

Data modeling
lm function will give linear regression
VIF is variance inflation factor
```{r}
# Building on train
new=new[,-9]
model=lm(CompressiveStrength1~., data=new)
summary(model)
VIF(model)
new1=new[,-c(5,6,7)]
model1=lm(CompressiveStrength1~., data=new1)
summary(model1)
VIF(model1)
```

Model validation
R^2 will determine performation of model in predicted value.
```{r}
test=read.csv('test.csv')
str(test)
test$CompressiveStrength1=test$CompressiveStrength^0.5
test=test[,-9]
pred=predict(model1,test)
sse=sum((pred-test$CompressiveStrength1)^2)
sst=sum((test$CompressiveStrength1-mean(test$CompressiveStrength1))^2)
e=(sse/sst)
rsquare=1-e
rsquare
```

```{r}
#test
plot(pred,test$CompressiveStrength1, main='test:Predicted vs Actual value', xlab = 'Predicted value',ylab = 'Actual value', col=c('blue','yellow'))
#train
plot(fitted.values(model1),new1$CompressiveStrength1, main='train: Predicted vs Actual value', xlab = 'Predicted value',ylab = 'Actual value', col=c('blue','yellow'))

#Model assumption
## Residual normality
plot(density(resid(model1)), main= 'Residual model', col='red')
skewness(resid(model1))
qqnorm(resid(model1))
qqline(resid(model1),col='red')
# Residual homoscedasticity
plot(fitted.values(model1),resid(model1),main='Fitted vs Residual value',xlab='fitted value',ylab='residual value', col=c('red','green'))

```

homoscedacit
```{r}
#Breuch pagan test for homoscedacity
lmtest::bptest(model1)

```

Residual liniarity
```{r}
#Residual liniarity

new1$resid=resid(model1)
plot(new$Cement,resid(model1),xlab = 'Cement (kg)',ylab = 'Residual', main = 'Cement Residual linearity',col=c('blue','orange'))
plot(new$Slag,resid(model1),xlab = 'Slag (kg)',ylab = 'Residual', main = 'Slag Residual linearity',col=c('red','purple'))
plot(new$Ash,resid(model1),xlab = 'Ash (kg)',ylab = 'Residual', main = 'Ash Residual linearity',col=c('blue','orange'))
plot(new$Water,resid(model1),xlab = 'Water (kg)',ylab = 'Residual', main = 'Water Residual linearity',col=c('green','purple'))
plot(new$Superplasticizer,resid(model1),xlab = 'Superplasticizer (kg)',ylab = 'Residual', main = 'Superplasticizer Residual linearity',col=c('brown','orange'))
plot(new$CoarseAggregate,resid(model1),xlab = 'CoarseAggregate (kg)',ylab = 'Residual', main = 'CoarseAggregate Residual linearity',col=c('red','yellow'))
plot(new$FineAggregate,resid(model1),xlab = 'FineAggregate (kg)',ylab = 'Residual', main = 'FineAggregate Residual linearity',col=c('blue','orange'))
plot(new$FineAggregate,resid(model1),xlab = 'FineAggregate (kg)',ylab = 'Residual', main = 'FineAggregate Residual linearity',col=c('red','purple'))

# Model validation
res = cv.lm(data=new, (form.lm =CompressiveStrength1~Cement+Slag+Ash+Water ),m=5,dots=FALSE,seed=29, plotit = c("Observed", "Residual"), legend.pos="topleft")


```

K fold summary
```{r}
#k fold summary
k=5
new$id <- sample(1:k, nrow(new), replace = TRUE)
list <- 1:k
for(i in 1:k){
  trainset <- subset(new,id %in% list[-i])
  testset <- subset(new,id %in% c(i))
  model <- lm(trainset$CompressiveStrength1~trainset$Cement+trainset$Slag+trainset$Ash+trainset$Water)
  predictions <- predict(model,testset[-c(4,5)])
  a <- summary(model)
   rmse <- sqrt(mean(model$residuals^2))
   print(model$coefficients)
   print(a$r.squared)
   print(rmse)
 }

# #Actual Vs. Fitted value
plot(fitted.values(model),new$CompressiveStrength1,col=c("red","blue"),main="Actual value vs. Fitted value")#Actual Vs. fitted
```

